---
title: "caseStudy3finalScript"
author: "JeanPhilipJuachon"
date: "10/24/2019"
output: html_document
---

```{r load philip defaults, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(ggfortify)
library(ggplot2)
library(ISLR)
library(leaps)
library(caret)
library(skimr)
library(corrplot)
library(glmnet)
library(ResourceSelection)
library(plotROC)
library(pROC)
library(ROSE)
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE, 
                      fig.align = 'center', fig.width = 5, fig.height = 5)
```

###loading the data
```{r}
data <- read.csv("training_data.csv", stringsAsFactors = TRUE)
testdata <- read.csv("test_data.csv")
```

###removal of columns that did not meet the conditions
```{r}
data <- select(data, -c(encounter_id, weight, glimepiride.pioglitazone, citoglipton, examide,medical_specialty, acetohexamide))
```

###cleansing and transformation (removal of "?", transformation of readmitted to binary)
```{r}
data$readmitted <- as.integer(ifelse(data$readmitted == "<30", 1, 0))
data$readmitted <- as.factor(data$readmitted)

as.factor(data$gender[data$gender == "Unknown/Invalid"] <- "Female")
data$gender <- factor(data$gender)

as.factor(data$race[data$race == "?"] <- "Other") 
data$race <- factor(data$race)

data$payer_code[data$payer_code == "?" | data$payer_code == "FR"| data$payer_code == "MP" | data$payer_code == "SI"] <- "OT"
data$payer_code <- factor(data$payer_code)

data <- data[data$diag_1 != "?",]
data <- data[data$diag_2 != "?",]
data <- data[data$diag_3 != "?",]
data$diag_1 <- factor(data$diag_1)
data$diag_2 <- factor(data$diag_2)
data$diag_3 <- factor(data$diag_3)

#factor the nominal values
data$admission_type_id <- as.factor(data$admission_type_id)
data$admission_source_id <- as.factor(data$admission_source_id)
data$discharge_disposition_id <- as.factor(data$discharge_disposition_id)
```

##Creation of Service Utilization
```{r}
data$service_utilization <- data$number_outpatient + data$number_emergency + data$number_inpatient
data <- select(data, -c(number_outpatient, number_emergency, number_inpatient))
```

#getting the count of diabetes medication
```{r}
data$metformin <- ifelse(data$metformin == "No",0,1)
data$repaglinide <- ifelse(data$repaglinide == "No",0,1)
data$nateglinide <- ifelse(data$nateglinide == "No",0,1)
data$chlorpropamide <- ifelse(data$chlorpropamide == "No",0,1)
data$glimepiride <- ifelse(data$glimepiride == "No",0,1)
data$glipizide <- ifelse(data$glipizide == "No",0,1)
data$glyburide <- ifelse(data$glyburide == "No",0,1)
data$tolbutamide <- ifelse(data$tolbutamide == "No",0,1)
data$pioglitazone <- ifelse(data$pioglitazone == "No",0,1)
data$rosiglitazone <- ifelse(data$rosiglitazone == "No",0,1)
data$acarbose <- ifelse(data$acarbose == "No",0,1)
data$miglitol <- ifelse(data$miglitol == "No",0,1)
data$troglitazone <- ifelse(data$troglitazone == "No",0,1)
data$tolazamide <- ifelse(data$tolazamide == "No",0,1)
data$glyburide.metformin <- ifelse(data$glyburide.metformin == "No",0,1)
data$metformin.rosiglitazone <- ifelse(data$metformin.rosiglitazone == "No",0,1)
data$metformin.pioglitazone <- ifelse(data$metformin.pioglitazone == "No",0,1)
data$glipizide.metformin <- ifelse(data$glipizide.metformin == "No",0,1)

data$count_of_diabetic_meds <- data$glipizide.metformin + data$metformin.pioglitazone + data$metformin.rosiglitazone+
  data$glyburide.metformin + data$tolazamide + data$troglitazone + data$miglitol + data$acarbose + data$rosiglitazone+
  data$pioglitazone + data$tolbutamide+ data$glyburide+data$glipizide+ data$glimepiride +
  data$chlorpropamide + data$nateglinide + data$repaglinide + data$metformin

data<-select(data, -c(19:32))
data <- select(data, -c(glyburide.metformin, metformin.pioglitazone, metformin.rosiglitazone
                        ,glipizide.metformin))
#data <- select(data, -c(troglitazone,tolazamide))
```

#Transformation/grouping of the diag levels to circulatory, respiratory, digestive, etc.
```{r}
data2 <- data
#diagnosis1
data2$diagnosis_group <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_group[data2$diag_1 %in% as.character(390:459) | data2$diag_1=='785'] <- "circulatory"
data2$diagnosis_group[data2$diag_1 %in% as.character(460:519) | data2$diag_1=='786'] <- "respiratory"
data2$diagnosis_group[data2$diag_1 %in% as.character(520:579) | data2$diag_1=='787'] <- "Digestive"
data2$diagnosis_group[substr(data2$diag_1,1,3) %in% as.character(250)] <- "Diabetes"
data2$diagnosis_group[data2$diag_1 %in% as.character(800:999)] <- "Injury"
data2$diagnosis_group[data2$diag_1 %in% as.character(710:739)] <- "Musculoskeletal"
data2$diagnosis_group[data2$diag_1 %in% as.character(580:629) | data2$diag_1=='788'] <- "Genitourinary"
data2$diagnosis_group[data2$diag_1 %in% as.character(140:239) | data2$diag_1 %in% as.character(790:799) | data2$diag_1=='780' | data2$diag_1 %in% as.character(240:249) |
                        data2$diag_1 %in% as.character(251:279) | data2$diag_1 %in%as.character(680:709) |
                        data2$diag_1 %in% as.character(001:139) | data2$diag_1=='781' |
                      data2$diag_1=='782' | data2$diag_1=='784'] <- "Neoplasms"
#diagnosis2
data2$diagnosis_2 <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_2[data2$diag_2 %in% as.character(390:459) | data2$diag_2=='785'] <- "circulatory"
data2$diagnosis_2[data2$diag_2 %in% as.character(460:519) | data2$diag_2=='786'] <- "respiratory"
data2$diagnosis_2[data2$diag_2 %in% as.character(520:579) | data2$diag_2=='787'] <- "Digestive"
data2$diagnosis_2[substr(data2$diag_2,1,3) %in% as.character(250)] <- "Diabetes"
data2$diagnosis_2[data2$diag_2 %in% as.character(800:999)] <- "Injury"
data2$diagnosis_2[data2$diag_2 %in% as.character(710:739)] <- "Musculoskeletal"
data2$diagnosis_2[data2$diag_2 %in% as.character(580:629) | data2$diag_2=='788'] <- "Genitourinary"
data2$diagnosis_2[data2$diag_2 %in% as.character(140:239) | data2$diag_2 %in% as.character(790:799) | data2$diag_1=='780' | data2$diag_2 %in% as.character(240:249) |
                        data2$diag_2 %in% as.character(251:279) | data2$diag_2 %in% as.character(680:709) |
                        data2$diag_2 %in% as.character(001:139) | data2$diag_2 =='781' |
                      data2$diag_2 =='782' | data2$diag_2=='784'] <- "Neoplasms"
#diagnosis3
data2$diagnosis_3 <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_3[data2$diag_3 %in% as.character(390:459) | data2$diag_3=='785'] <- "circulatory"
data2$diagnosis_3[data2$diag_3 %in% as.character(460:519) | data2$diag_3=='786'] <- "respiratory"
data2$diagnosis_3[data2$diag_3 %in% as.character(520:579) | data2$diag_3=='787'] <- "Digestive"
data2$diagnosis_3[substr(data2$diag_3,1,3) %in% as.character(250)] <- "Diabetes"
data2$diagnosis_3[data2$diag_3 %in% as.character(800:999)] <- "Injury"
data2$diagnosis_3[data2$diag_3 %in% as.character(710:739)] <- "Musculoskeletal"
data2$diagnosis_3[data2$diag_3 %in% as.character(580:629) | data2$diag_3=='788'] <- "Genitourinary"
data2$diagnosis_3[data2$diag_3 %in% as.character(140:239) | data2$diag_3 %in% as.character(790:799) | data2$diag_1=='780' | data2$diag_3 %in% as.character(240:249) |
                        data2$diag_3 %in% as.character(251:279) | data2$diag_3 %in%as.character(680:709) |
                        data2$diag_3 %in% as.character(001:139) | data2$diag_3=='781' |
                      data2$diag_3=='782' | data2$diag_3=='784'] <- "Neoplasms"

data2 <- select(data2, -diag_1, -diag_2, -diag_3)
data2 <- rename(data2, diag1 = diagnosis_group, diag2=diagnosis_2, diag3 = diagnosis_3)
```

##transforming admission_type_id
```{r}
data2$admission_type <- factor(rep("other", nrow(data2)), ordered = F,
                                  levels = c("Non-elective", "Elective","other"))
data2$admission_type[data2$admission_type_id == 2 | data2$admission_type_id == 1 |data2$admission_type_id == 7] <- "Non-elective"
data2$admission_type[data2$admission_type_id == 3 | data2$admission_type_id == 4] <- "Elective"

data2 <- select(data2, -admission_type_id)
```

#removing all those who are dead/hospice/expired
#note that this code can either be used or not
#running this code will give you a GOF of 0.058, not running will give you a gof of 0.16
```{r}
#data2 <- data2[data2$discharge_disposition_id != '11',]
#data2 <- data2[data2$discharge_disposition_id != '13',]
#data2 <- data2[data2$discharge_disposition_id != '14',]
#data2 <- data2[data2$discharge_disposition_id != '19',]
#data2 <- data2[data2$discharge_disposition_id != '20',]
#data2 <- data2[data2$discharge_disposition_id != '21',]
```

#creating new factors for the discharge_disposition
```{r}
data2$discharge_disposition <- factor(rep("Transferred", nrow(data2)), ordered = F,
                                      levels = c("Home","Transferred", "Left_AMA"))
data2$discharge_disposition[data2$discharge_disposition_id== 1 | data2$discharge_disposition_id== 6 |data2$discharge_disposition_id== 8] <- "Home"
data2$discharge_disposition[data2$discharge_disposition_id== 7] <- "Left_AMA"

data2 <- select(data2, -discharge_disposition_id)
```

#new factors for admission source
```{r}
data2$admission_source <- factor(rep("Others", nrow(data2)), ordered=  F,
                                 levels = c("Referrals", "Transferred", "Others")
                                 )
data2$admission_source[data2$admission_source_id ==1 | data2$admission_source_id ==2 |data2$admission_source_id ==3 ] <- "Referrals"
data2$admission_source[data2$admission_source_id ==4 | data2$admission_source_id ==5 |data2$admission_source_id ==6 |data2$admission_source_id ==10|data2$admission_source_id ==18|data2$admission_source_id ==22 |data2$admission_source_id ==25 |data2$admission_source_id ==26 ] <- "Transferred"

data2 <- select(data2, -admission_source_id)
#data2$number_of_med_change <- as.integer(data2$number_of_med_change)
```

#relevel of factors
```{r}
data2<- data2 %>% 
  within(diag1 <- relevel(diag1, ref = "other")) %>% 
  within(diag2 <- relevel(diag2, ref = "other")) %>% 
  within(diag3 <- relevel(diag3, ref = "other"))
```

#changing the base level of max_glu_serum A1Cresult, insulin, change, and admission source
```{r}
data2 <- data2 %>% 
  within(max_glu_serum <- relevel(max_glu_serum, ref = "None")) %>% 
  within(A1Cresult <- relevel(A1Cresult, ref = "None")) %>% 
  within(insulin <- relevel(insulin, ref = "No")) %>% 
  within(change <- relevel(change, ref = "No")) %>% 
  within(admission_source <- relevel(admission_source, ref = "Others"))
```

#goodness of fit
```{r}
gof1 <- glm(readmitted~race+payer_code+number_diagnoses+discharge_disposition+num_lab_procedures+
                max_glu_serum+diag1+diag2+diag3+admission_source+age+num_procedures+A1Cresult+
              service_utilization+
                num_medications+change+count_of_diabetic_meds+admission_type + insulin
            ,data = data2, family = binomial())

pchisq(gof1$deviance, gof1$df.residual, lower.tail = FALSE)
hoslem.test(gof1$y, fitted(gof1))
```
```{r}
summary(gof1)
```

#model using all predictors
```{r}
inference <- glm(readmitted~.-patient_nbr, data = data2, family = binomial())
summary(inference)
```



###preparation for prediction, training and test split of the training data
```{r}
training <- createDataPartition(y = data2$readmitted, p = .70, list = FALSE)
trainingset <- data2[training,]
validationset <- data2[-training,]
oversampledTrain <- ovun.sample(readmitted ~ . ,data = trainingset, method = "over", N = 39465)$data
undersampledTrain <- ovun.sample(readmitted ~ . ,data = trainingset, method = "under", N = 38465)$data
bothsampledTrain <- ovun.sample(readmitted ~ . ,data = trainingset, method = "both", N = 38465)$data
```

#models with good fit
```{r}
model1 <- glm(readmitted~race+payer_code+number_diagnoses+discharge_disposition+num_lab_procedures+
                max_glu_serum+diag1+diag2+diag3+admission_source+age+num_procedures+A1Cresult+
              service_utilization+
                num_medications+change+count_of_diabetic_meds+admission_type + insulin
            ,data = oversampledTrain, family = binomial())
```

#trial prediction using model 2
```{r}
model2 <- glm(readmitted~ race + payer_code + number_diagnoses+discharge_disposition + num_lab_procedures + 
               diag1+diag2+diag3+num_procedures + service_utilization + num_medications + insulin
            ,data = bothsampledTrain, family = "binomial")

preds2 <- predict(model2, validationset, type = "response")
preds2 <- ifelse(preds2 > .6, 1, 0)
results2 <- table(validationset$readmitted,preds2)
```


```{r}


#model3 <- glm(readmitted ~race+gender+age+admission_type+discharge_disposition+admission_source+time_in_hospital+num_lab_procedures+num_procedures+num_medications+service_utilization+diag1+diag2+diag3+number_diagnoses+max_glu_serum+A1Cresult+insulin
 #             ,data = bothsampledTrain, family = "binomial")

#model4 <- glm(readmitted~race + age+time_in_hospital + payer_code + num_lab_procedures+num_procedures+num_medications+number_diagnoses+A1Cresult + insulin+change+diabetesMed+service_utilization+diag1+diag2+diag3+discharge_disposition+admission_source
  #            , data = trainingset, family = "binomial")
```

#getting the auc
```{r}
real_val = as.numeric(data2$readmitted)
preds = predict(model2, data2, type = "response")
data.frame(real_val = real_val, preds = preds) %>% 
ggplot(aes(d=real_val, m=preds)) +
  geom_roc(n.cuts = 6, labelround = 4) +
  geom_abline(intercept = 0, slope = 1)
roc1 <- roc(real_val, preds)
auc(roc1)
```

```{r}
gof1a <- glm(readmitted~ race + payer_code + number_diagnoses+discharge_disposition + num_lab_procedures + 
               diag1+diag2+diag3+num_procedures +service_utilization + num_medications + insulin
             ,data = data2, family = "binomial")
pchisq(gof1a$deviance, gof1a$df.residual, lower.tail = FALSE)
hoslem.test(gof1a$y, fitted(gof1a))
summary(gof1a)
```


#This portion is used to fill up the readmission probabilities of the test data set
#In order to run properly, change the file name of the test data set to "data" and then run all the codes
#that is related to cleansing only. This is in order to follow the cleansing steps done for the training data.
#Improper cleansing of the test data may result to failure in the model as there might be levels/variables that are not included in the training data that might show up in the test set.

#predicting using hybridsampled data model
```{r}
pred_prob <- predict(model2, data2, type = "response")
#preds <- as.factor(ifelse(preds > 5 , 1,0))

#confusionMatrix(preds, validationset$readmitted, positive = "1")
```
#in order to predict the test_data set, just rename the test data to "data" then run all the cleansing done on the #training data. 
```{r}
output <- cbind(data2, pred_prob)
output <- select(output, patient_nbr, pred_prob)
write.csv(output, 'C:\\Users\\jjuachon\\Desktop\\juachon_jean.csv')
```













